---
title: "descriptives"
author: "Anna Fuchs, Niklas Ippisch"
format: html
---

# Preperation
```{r}
library(tidyverse)
library(ggpubr)
relevant <- read_rds(file = "data/relevant_recoded.R")
```


# Descriptives

```{r}
relevant |> 
  filter(w2eq1 == 3 & eisced != 55) |> 
  ggplot(aes(x = hinctnta)) +
  geom_bar()
```

```{r}
relevant |> 
  filter(is.na(w2eq13)) |> 
  ggplot(aes(x = eduyrs)) +
  geom_bar()
```


## Dependent variable: In favour or against increasing spending on social assistance on the poor (w2eq13)
```{r}
relevant |> 
  filter(!is.na(w2eq13)) |> 
  ggplot(aes(x = w2eq13)) +
  geom_bar()
```

```{r}
relevant |> 
  filter(!is.na(w2eq11)) |> 
  ggplot(aes(x = w2eq11)) +
  geom_bar()
```

Influence of control variables

```{r}
# Control variable: Age
eq13_byage <- relevant |> 
  filter(!is.na(w2eq13) & !is.na(age)) |> 
  group_by(age) |> 
  summarize(mean = mean(w2eq13, na.rm = TRUE)) |> 
  ggplot(aes(x = age, y = mean)) +
  geom_point()+
  ylim(1, 5)

# Control variable: Gender
eq13_bygendr <- relevant |> 
  filter(!is.na(w2eq13) & !is.na(female)) |> 
  ggplot(aes(x = w2eq13, fill = factor(female))) +
  geom_bar(position = "dodge")

# Control variable: Income percentile
eq13_byincome <- relevant |> 
  filter(!is.na(w2eq13) & !is.na(hinctnta)) |> 
  group_by(hinctnta) |> 
  summarize(mean = mean(w2eq13, na.rm = TRUE)) |> 
  ggplot(aes(x = hinctnta, y = mean)) +
  geom_point() +
  ylim(1, 5)

# Control variable: Education years
eq13_byedu <- relevant |> 
  filter(!is.na(w2eq13) & !is.na(eisced) & eisced != 55) |> 
  group_by(eisced) |> 
  summarize(mean = mean(w2eq13, na.rm = TRUE)) |> 
  ggplot(aes(x = eisced, y = mean)) +
  geom_point()+
  ylim(1, 5)
ggarrange(eq13_byage, eq13_bygendr, eq13_byincome, eq13_byedu, nrow = 2, ncol = 2)
```

Influence of possible other variables

```{r}
# Possible other influencing variable: Meritocratic beliefs (hard work important for success, w1eq10)
eq13_bymerit <- relevant |> 
  filter(!is.na(w2eq13) & !is.na(w1eq10)) |> 
  group_by(w1eq10) |> 
  summarize(mean = mean(w2eq13, na.rm = TRUE)) |> 
  ggplot(aes(x = w1eq10, y = mean)) +
  geom_point() +
  ylim(1, 5)

# Possible other influencing variable: tension between rich and poor people, w1eq3
eq13_bytension <- relevant |> 
  filter(!is.na(w2eq13) & !is.na(w1eq3)) |> 
  group_by(w1eq3) |> 
  summarize(mean = mean(w2eq13, na.rm = TRUE)) |> 
  ggplot(aes(x = w1eq3, y = mean)) +
  geom_point() +
  ylim(1, 5)

# Possible other influencing variable: reducing differences is in government's responsibility, w1eq15
eq13_bygov <- relevant |> 
  filter(!is.na(w2eq13) & !is.na(w1eq15)) |> 
  group_by(w1eq15) |> 
  summarize(mean = mean(w2eq13, na.rm = TRUE)) |> 
  ggplot(aes(x = w1eq15, y = mean)) +
  geom_point() +
  ylim(1, 5)

# Possible other influencing variable: how angry about wealth(!) differences, w1eq8
eq13_byangry <- relevant |> 
  filter(!is.na(w2eq13) & !is.na(w1eq8)) |> 
  group_by(w1eq8) |> 
  summarize(mean = mean(w2eq13, na.rm = TRUE)) |> 
  ggplot(aes(x = w1eq8, y = mean)) +
  geom_point() +
  ylim(1, 5)

ggarrange(eq13_bymerit, eq13_bytension, eq13_bygov, eq13_byangry, nrow = 2, ncol = 2)

```


## Independent variable: Income inequality in country increased or decreased past 5 years (w2eq1)
```{r}
relevant |> 
  filter(!is.na(w2eq1)) |> 
  ggplot(aes(x = w2eq1)) +
  geom_bar()
```

```{r}
# Control variable: Age
eq1_byage <- relevant |> 
  filter(!is.na(w2eq1) & !is.na(age)) |> 
  group_by(age) |> 
  summarize(mean = mean(w2eq1, na.rm = TRUE)) |> 
  ggplot(aes(x = age, y = mean)) +
  geom_point() +
  ylim(1, 5)

# Control variable: Gender
eq1_bygendr <- relevant |> 
  filter(!is.na(w2eq1) & !is.na(female)) |> 
  ggplot(aes(x = w2eq1, fill = factor(female))) +
  geom_bar(position = "dodge")

# Control variable: Income decentile
eq1_byincome <- relevant |> 
  filter(!is.na(w2eq1) & !is.na(hinctnta)) |> 
  group_by(hinctnta) |> 
  summarize(mean = mean(w2eq1, na.rm = TRUE)) |> 
  ggplot(aes(x = hinctnta, y = mean)) +
  geom_point() +
  ylim(1, 5)

# Control variable: Education years
eq1_byedu <- relevant |> 
  filter(!is.na(w2eq1) & !is.na(eduyrs)) |> 
  group_by(eduyrs) |> 
  summarize(mean = mean(w2eq1, na.rm = TRUE)) |> 
  ggplot(aes(x = eduyrs, y = mean)) +
  geom_point() +
  ylim(1, 5)
ggarrange(eq1_byage, eq1_bygendr, eq1_byincome, eq1_byedu, nrow = 2, ncol = 2)
```

# Correlation analyses

Correlation between the support on spending on social assistance and perceived inequality

From here it's just playing around

```{r}
relevant |> 
  filter(!is.na(w2eq13) & !is.na(w2eq1)) |> 
  group_by(w2eq1) |> 
  summarize(mean = mean(w2eq13, na.rm = TRUE)) |> 
  ggplot(aes(x = w2eq1, y = mean)) +
  geom_point() +
  ylim(1, 5)
```

```{r}
cor1 <- relevant |> 
  filter(!is.na(w2eq13) & !is.na(w2eq1))

cor(cor1$w2eq13, cor1$w2eq1, method = "spearman") # -0.117

m1 <- lm(w2eq13 ~ w2eq1, data = relevant)
summary(m1)

```

```{r}

m2 <- lm(w2eq13 ~ w2eq1 + w1eq10 + w1eq3, data = relevant)
summary(m2)
```

```{r}
m3 <- lm(w2eq13 ~ w2eq1 + age + gndr + hinctnta + eduyrs, data = relevant)
summary(m3)
```

```{r}
relevant1 <- relevant |> 
  filter(w2eq13 != 3 & w2eq1 != 3) |> 
  mutate(policy_bin = ifelse(w2eq13 == 1 | w2eq13 == 2, 1, 0),
         income_bin = ifelse(w2eq1 == 5 | w2eq1 == 4, 1, 0))
table(relevant1$policy_bin, relevant1$income_bin)
chisq.test(relevant1$policy_bin, relevant1$income_bin)
```

```{r}
m4 <- glm(policy_bin ~ income_bin, data = relevant1, family = "binomial")
summary(m4)

m5 <- glm(policy_bin ~ income_bin + age + gndr + hinctnta + eduyrs, data = relevant1, family = "binomial")
summary(m5)
```

